---
title: "Data analysis with R final exam 2022"
date: '2022-07-04'
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r message = F, echo = F}
library(tidyverse)
```

## 1. 

```{r}
tmp <- c(4, 6, 3)
```

Create the vectors

(a) $(4, 6, 3, 4, 6, 3, \dots, 4, 6, 3)$ where there are 10 occurrences of 4.
    ```{r}
    rep(c(4,6,3), times=10)
    ```
(b) $(4, 4, \dots, 4, 6, 6, \dots, 6, 3, 3, \dots, 3)$ where there are 10 occurrences of 4, 20 occurrences of 6 and 30 occurrences of 3.
    ```{r}
    cat(rep(4, each=10), rep(6, each=20), rep(3, each=30))
    ```

## 2. 
Execute the following lines which create two vectors of random integers which are chosen with replacement from the integers $0, 1, \dots , 999$. Both vectors have length 250. 

```{r}
xVec <- sample(0:999, 250, replace=T)
yVec <- sample(0:999, 250, replace=T)
```

(a) Create the vector $(y_2 - x_1, \cdots, y_n - x_{n-1}).$
    ```{r}
    tmp <- yVec[-1]-xVec[-250]
    ```
(b) Pick out the values in yVec which are $> 600$.
    ```{r}
    tmp <- yVec[yVec > 600]
    ```
(c) What are the index positions in yVec of the values which are $> 600$? 
    ```{r}
    which(yVec > 600)
    ```
(d) Sort the numbers in the vector xVec in the order of increasing values in yVec. 
    
(e) Pick out the elements in yVec at index positions $1, 4, 7, 10, 13, \cdots$
    ```{r}
    idx <- seq(1,250, by=3)
    tmp <- yVec[idx]
    ```

## 3.
By using the function cumprod and other functions to calculate:

\[ 1 + \frac{2}{3} + \left(\frac{2}{3}\frac{4}{5}\right) + \left(\frac{2}{3}\frac{4}{5}\frac{6}{7}\right) + \cdots + \left(\frac{2}{3}\frac{4}{5}\cdots\frac{38}{39}\right)
\]

```{r}
x <- seq(2,38, by=2) / seq(3,39, by=2)
result <- 1 + sum(cumprod(x))
result
```

## 4. 
For this problem we'll use the (built-in) dataset state.x77.
```{r}
data(state)
state.x77 <- as.data.frame(state.x77)
head(state.x77)
```

a. Find out how many states have an income of less than 4300.
    ```{r}
    length(state.x77$Income < 4300)
    ```

b.	Find out which is the state with the highest income.
    ```{r}
    rownames(state.x77[which(state.x77$Income==max(state.x77$Income)),])
    ```

c.	Add a variable to the data frame which should categorize the level of illiteracy: $[0,1)$ is low, $[1,2)$ is some, $[2, \infty)$ is high.
    ```{r}
    library(dplyr)
    state.x77 <- state.x77 %>% mutate(Level = 
                ifelse(Illiteracy<1, "low",
                    ifelse(Illiteracy<2, "some", "high")))
    head(state.x77)
    ```

d.	Find out which state with low illiteracy, has the highest income, and what that income is.
    ```{r}
    state.tmp <- state.x77 %>% filter(Level=="low")
    max(state.tmp$Income)
    rownames(state.tmp[which(state.tmp$Income==max(state.tmp$Income)),])
    ```

## 5. 
Simulate 1,000 observations from $(X_1, X_2)$ which follow the uniform distribution over the square $[0, 1] \times [0, 1]$.
```{r}
x1 <- runif(1000, 0,1)
x2 <- runif(1000, 0,1)
```
a. Get an approximation of the probability that the distance between $(X_1, X_2)$ and the nearest edge is less than 0.25.
    ```{r}
    length(x1[x1<0.25 | 1-x1<0.25 | x2<0.25 | 1-x2<0.25])/1000
    ```

b. The same question for the distance to the nearest vertex.
    ```{r}
    length(x1[x1^2+x2^2<0.25^2 | (1-x1)^2+x2^2<0.25^2 |
        x1^2+(1-x2)^2<0.25^2 | (1-x1)^2+(1-x2)^2<0.25^2])/1000
    ```

## 6.
A discrete random variable $X$ has probability mass function 

```{=latex}
\begin{center}\begin{tabular}{lrrrrr} $x$ & 0 & 1 & 2 & 3 & 4 \\\hline $p(x)$ & 0.1 & 0.2 & 0.2 & 0.2 & 0.3 \end{tabular}\end{center}
```

Generate a random sample of size 1000 from the distribution of $X$ using the R `sample()` function. Construct a relative frequency table and compare the empirical with the theoretical probabilities. 
```{r}
x <- c(0, 1,1, 2,2, 3,3, 4,4,4)
samp <- sample(x, size=1000, replace=T)
res <- data.frame(
    x_equal_0 = c(0.1, length(samp[samp==0])/1000),
    x_equal_1 = c(0.2, length(samp[samp==1])/1000),
    x_equal_2 = c(0.2, length(samp[samp==2])/1000),
    x_equal_3 = c(0.2, length(samp[samp==3])/1000),
    x_equal_4 = c(0.3, length(samp[samp==4])/1000)
)
rownames(res) <- c("theoretical_Pr", "empirical_Pr")
res
```


## 7. 
Mortality rates per 100,000 from male suicides for a number of age groups and a number of countries are given in the following data frame. 
```{r}
suicrates <- tibble(Country = c('Canada', 'Israel', 'Japan', 'Austria', 'France', 'Germany',
'Hungary', 'Italy', 'Netherlands', 'Poland', 'Spain', 'Sweden', 'Switzerland', 'UK', 'USA'), 
Age25.34 = c(22,  9, 22, 29, 16, 28, 48,  7,  8, 26,  4, 28, 22, 10, 20), 
Age35.44 = c(27, 19, 19, 40, 25, 35, 65,  8, 11, 29,  7, 41, 34, 13, 22), 
Age45.54 = c(31, 10, 21, 52, 36, 41, 84, 11, 18, 36, 10, 46, 41, 15, 28), 
Age55.64 = c(34, 14, 31, 53, 47, 49, 81, 18, 20, 32, 16, 51, 50, 17, 33), 
Age65.74 = c(24, 27, 49, 69, 56, 52, 107, 27, 28, 28, 22, 35, 51, 22, 37))
```

a. Transform `suicrates` into *long* form.
    ```{r}
    library(reshape)
    longdata <- melt(as.data.frame(suicrates), id="Country")
    longdata
    ```

b. Construct side-by-side box plots for the data from different age groups, and comment on what the graphic tells us about the data. 
    ```{r}
    library(ggplot2)
    ggplot(longdata, aes(x=variable, y=value, color=variable)) +
        geom_boxplot()
    ```
    As seen in the image, the lowest suicide rate is in the 25-34 age group and the highest in the 65-74 age group.
    As the age increases, the suicide rate is gradually increasing.

## 8. 
The steam data in the `MASS` package has a nonlinear regression model, 
\[P = \alpha \exp\left\{\frac{\beta t}{\gamma + t}\right\} + \varepsilon \]

Fit the  model with `nls()` function and find the fitted values, using initial value $\alpha = 5, \beta = 20, \gamma = 200$. Plot them with the data points on the original scale.

```{r message = F}
library(MASS)
data(steam)
f <- function(t, a,b,g){
    return (a*exp(b*t/(g+t)))
}
fitted.model <- nls(Press~f(Temp, a,b,g), 
        data=steam, start=list(a=5,b=20,g=200))
summary(fitted.model)
steam <- steam %>% mutate(FittedModel = 
                predict(fitted.model, steam))
ggplot(steam) + geom_point(aes(Temp, Press, color="red"))+
        geom_line(aes(Temp, Press, color="red")) +
        geom_point(aes(Temp, FittedModel, color="blue"))+
        geom_line(aes(Temp, FittedModel, color="blue"))
```
"red" is Press, while "blue" is FittedModel.