---
title: "Homework 3"
author: 'zhang zhuohan'
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

**Background**: ......
```{r}
library(ggplot2)
library(dplyr)
gmp <- read.table("../data/gmp.dat")
gmp$pop <- round(gmp$gmp/gmp$pcgmp)
```

1. First, plot the data as in lecture, with per capita GMP on the y-axis and population on the x-axis. 
   Add the curve function with the default values provided in lecture. Add two more curves corresponding to $a=0.1$ and $a=0.15$;
   use the `col` option to give each curve a different color (of your choice).
    ```{r}
    gmp %>% ggplot() + geom_point(aes(pop, pcgmp)) + 
        labs(x = "Population", y = "Per-Capita Economic Output ($/person-year)",
            title = "US Metropolitan Areas, 2006") +
        scale_x_log10() + 
        geom_line(aes(pop, 6611*pop^(0.125), color="blue"), size=1.5) +
        geom_line(aes(pop, 6611*pop^(0.1), color="red"), size=1.5) +
        geom_line(aes(pop, 6611*pop^(0.15), color="yellow"), size=1.5) +
        scale_colour_manual(name = "value of a",
                            values=c("yellow"="yellow", "blue"="blue", "red"="red"),
                            breaks=c("yellow", "blue", "red"),
                            labels=c("a=0.1", "a=0.125", "a=0.15"))
    ```

2. Write a function, called `mse()`, which calculates the mean squared error of the model on a given data set.
   `mse()` should take three arguments: a numeric vector of length two, the first component standing for $y_0$
   and the second for $a$; a numerical vector containing the values of $N$; and a numerical vector containing 
   the values of $Y$. The function should return a single numerical value. The latter two arguments should have
   as the default values the columns `pop` and `pcgmp` (respectively) from the `gmp` data frame from lecture.
   Your function may not use `for()` or any other loop. Check that, with the default data, you get the following values.
    ```{r}
    mse <- function(coeff, N=gmp$pop, Y=gmp$pcgmp){
        return (mean((Y - coeff[1]*N^coeff[2])^2))
    }
    mse(c(6611,0.15))
    mse(c(5000,0.10))
    ```

3. **R** has several built-in functions for optimization, which we will meet as we go through the course.
   One of the simplest is `nlm()`, or non-linear minimization. `nlm()` takes two required arguments:
   a function, and a starting value for that function. Run `nlm()` three times with 
   your function `mse()` and three starting value pairs for $y_0$ and $a$ as in
    ```{r, eval = FALSE}
    nlm(mse, c(y0=6611,a=1/8))
    ```
    What do the quantities **minimum** and **estimate** represent? What values does it return for these?
    ```{r}
    nlm(mse, c(y0=6611,a=0.1))
    nlm(mse, c(y0=6611,a=0.125))
    nlm(mse, c(y0=6611,a=0.15))
    ```
    + **minimum** represents the value of the estimated minimum of the function.
    + **estimate** represents the the point at which the minimum value of the function is obtained.

4. Using `nlm()`, and the `mse()` function you wrote, write a function, `plm()`, which estimates the 
   parameters y0 and a of the model by minimizing the mean squared error. It should take the following arguments:
   an initial guess for $y_0$; an initial guess for $a$; a vector containing the $N$ values; a vector containing 
   the $Y$ values. All arguments except the initial guesses should have suitable default values. It should return a list 
   with the following components: the final guess for $y_0$; the final guess for $a$; the final value of the MSE. 
   Your function must call those you wrote in earlier questions (it should not repeat their code), and the 
   appropriate arguments to `plm()` should be passed on to them.
    ```{r}
    plm <- function(coeff, N = gmp$pop, Y = gmp$pcgmp){
        res <- nlm(mse, c(coeff[1],coeff[2]), N, Y)
        return (list(final_guess=c(res$estimate[1],res$estimate[2]),
                     MSE=res$minimum))
    }
    ```
   What parameter estimate do you get when starting from $y_0=6611$ and $a=0.15$? From $y_0=5000$ and $a=0.10$? 
   If these are not the same, why do they differ? Which estimate has the lower MSE?
    ```{r}
    plm(c(6611,0.15))
    plm(c(5000,0.10))
    ```
    The difference may be caused because there are multiple local optimal values.

5. *Convince yourself the jackknife can work.*
  + a. Calculate the mean per-capita GMP across cities, and the standard error of this mean, using the built-in functions `mean()` and `sd()`,
       and the formula for the standard error of the mean you learned in your intro. stats. class (or looked up on Wikipedia…).
        ```{r}
        mean(gmp$pcgmp) # mean
        sd(gmp$pcgmp)/sqrt(nrow(gmp)) # standard error
        ```
        The standard error is calculated as $\sqrt{\frac{1}{n(n-1)}\sum_{i=1}^{n}(x_i-\bar{x})^2}$.
  + b. Write a function which takes in an integer $i$, and calculate the mean per-capita GMP for every city except city number $i$.
        ```{r}
        mean.jackknife <- function(i, Y=gmp$pcgmp){
            return (mean(Y[-i]))
        }
        ```
  + c. Using this function, create a vector, `jackknifed.means`, which has the mean per-capita GMP where every city is held out in turn. (You may use a `for` loop or `sapply()`.)
        ```{r}
        n <- nrow(gmp)
        jackknifed.means <- sapply(1:n, mean.jackknife)
        ```
  + d. Using the vector `jackknifed.means`, calculate the jack-knife approximation to the standard error of the mean. How well does it match your answer from part (a)?
        ```{r}
        sqrt(((n-1)^2/n)*var(jackknifed.means))
        ```
        It's actually the same as the answer in part (a).

6. Write a function, `plm.jackknife()`, to calculate jackknife standard errors for the parameters $y_0$ and $a$.
   It should take the same arguments as `plm()`, and return standard errors for both parameters. This function should call your `plm()` function repeatedly. 
   What standard errors do you get for the two parameters?
    ```{r}
    plm.jackknife <- function(coeff, N=gmp$pop, Y=gmp$pcgmp){
        n <- length(N)
        pre.jackknife <- function(i){
            return(plm(coeff, N[-i], Y[-i])$final_guess)
        }
        jackknife.means <- sapply(1:n, pre.jackknife)
        result <- data.frame(
            y0.se = sqrt(((n-1)^2/n)*var(jackknife.means[1,])),
            a.se  = sqrt(((n-1)^2/n)*var(jackknife.means[2,]))
        )
        return(result)
    }
    plm.jackknife(c(6611, 0.125))
    ```

7. The file `gmp-2013.dat` contains measurements for for 2013. Load it, and use `plm()` and `plm.jackknife` to estimate the parameters of the model for 2013, and their standard errors. Have the parameters of the model changed significantly?
    ```{r}
    gmp2013 <- read.table("../data/gmp-2013.dat", header=T)
    gmp2013$pop <- round(gmp2013$gmp/gmp2013$pcgmp)
    plm(c(6611, 0.125), gmp2013$pop, gmp2013$pcgmp)
    plm.jackknife(c(6611, 0.125), gmp2013$pop, gmp2013$pcgmp)
    ```
    The parameters of the model didn't change significantly.